Lonely and online: an investigation of social media use and social isolation in China

### 2024-08-19 tandf/Response to reviewers second revision.qmd

---
title: "Response to reviewers - second revision"
format: pdf
---

## Reviewer 2

*1. Timeliness of Data: The research is based on data collected in 2015, a time when the prevalence of mobile internet and social media platforms might differ significantly from the present landscape. The authors utilized a survey conducted from November 25 to December 2, 2015, as mentioned in the document: "We employed Qualtrics to collect the data...from November 25 to December 2, 2015." Given the rapid evolution of digital communication, the applicability of findings based on a decade-old dataset should be reconsidered.*

*2. Mismatch Between Literature Review and Data Collection: The literature review incorporates studies from around 2020, such as the reference to "Valkenburg et al., 2022," yet the data used for analysis is from 2015. This discrepancy may affect the validity of conclusions drawn from contemporary theoretical perspectives when applied to outdated data.*

We are happy to revise our literature review section in ways that the reviewer feels would strengthen the paper but we are a bit unsure of how to proceed given two previous comments by reviewers that we add to the literature review section. While we acknowledge that the previous recommendation and this recommendation are not precisely mutually exclusive, it does suggest that if we remove some references here we may be working against the recommendations of the previous reviewers.

Suggestions as to how to move forward would be welcome here.

> Previous reviewer 1 comment: "It is suggested to examine the related theories to answer, why the phenomenon of social media use in Mainland China may be different from others."

> Previous reviewer 2 comment: "Nothing in the literature review about social media addiction. I suggest to see: Social isolation, social support and their relationship with smartphone addiction; examples: The relationship between addiction to smartphone usage and depression among adults: A crosssectional study, Smartphone addiction among university students in the light of some variables, and Examining the effects of motives and gender differences on smartphone addiction."

*3. Rationale for Variable Selection: The inclusion of variables like "Chinese Communist Party membership (CCP)" is noted, as stated: "For those we begin by fitting an additive model of our index of social isolationism as function of each of our primary independent variables...CCP Member." However, the introduction does not discuss the relevance of this variable, leaving unclear its contribution to the study.*

A reasonable point. We have added some discussion of the reason this variable is included in the final paragraph of the Data, Measurement and Description section.

*4. Depth of Discussion: The discussion section lacks depth, particularly in expanding upon the implications of the results. For example, the document mentions: "our results cast some doubt on theories that suggest it is the negative content..." but does not elaborate on how these findings challenge existing theories.*

Fair point. We have substantially increased discussion of the implications and linked it more explicitly back to the literature discussed in the theory section.

### 2024-08-19 tandf/Social Isolation in China second revision.qmd
Review 2 update

-   **H$1_a$**: social media use should have a positive connection to social isolation due to strengthened government control over online content

-   **H$1_b$**: social media use should have a negative connection to social isolation due to the dominance of influencer culture in China, moderated by age.

-   **H$2$**: those who are able to use social media to make new friends will have lower social isolation scores than those who do not

-   **H$3$:** the urban/rural divide will produce differential impacts in the above hypotheses on the relationship between social media use and social isolation

@@

We measured general social media use with an index based on the following three items: 1) About how many hours a day would you estimate you spend using only social media? Social media means applications like Weibo, QQ, Renren, Kaixin001, Douban, WeChat or other sites and services that allow users to interact with each other. (0-1, 1-2, 2-3, 3-4, 4-5, 5-6, 6-7, 7-8, 8-9, or More than 9), 2) Do you check email, read websites, and use social media (social media means applications like Weibo, QQ, Renren, Kaixin001, Douban, WeChat or other sites and services that allow users to interact with each other) more than you did five years ago? (Yes, No), and 3) How often do you read news stories about political events that have been posted on social media (social media means applications like Weibo, QQ, Renren, Kaixin001, Douban, WeChat or other sites and services that allow users to interact with each other)? (More than once a day, Everyday, Three-to-five days per week, One-to-two days per week, Less often, Never). Each was recoded so that higher values represented more social media use, rescaled to range from 0 through 1, added together, and then again, rescaled to range from 0 through 1 maintaining all original intervals.The Chronbach's alpha was relatively low for these three items ($\alpha = 0.31$), but we proceeded with the construction of this index because, first, the face validity is high. Simply, those who claim to use social media generally more often, claim to read political news more often, and claim to use it more than they did in the past, are certainly likely to be more frequent social media users than those who do not do all these things.Second, we confirmed that these three items were all positively correlated with each other (p < 0.0001 for all pairwise relationships). The distribution of this index is presented in @fig-sm-use-dist. The distribution is relatively normal and tightly centered around the mean 0.63 with a standard deviation of 0.17, but there is a left skew suggesting that a portion of the population is not that social media active, but clearly the bulk of the population are.

@@

Our modeling strategy that follows is straightforward. We begin by modeling each of our internet communication independent variables as function of socio-economic status (SES), gender, Chinese Communist Party membership (CCP), and age to create an individual profile of each type of digital user (see the Online Appendix for the operationalization of these variables). This provides some context for the models that follow which are intended to test our primary hypotheses, that is that internet communication can deter social isolation and that this is particularly true for those who do not have the convenience of the high social interaction opportunity provided by urban living. For those we begin by fitting an additive model of our index of social isolationism as function of each of our primary independent variables while also controlling for the same variables included in our profile models. Finally, we refit three models with the same specification but introduce an interactive term between each of our internet communication indicators, respectively, with our measure of urbanicity. This allows us to test whether the observed additive effects are stronger for those living outside cities.

@@

In terms of our original hypotheses, we find support for the idea that China's social media environment is not particularly unique with respect to generating feelings of social isolation (**H1b**). Whether or not this is due to the influencer culture or other factors, we cannot say with our data, however age is a consistently negative predictor of social isolation even when controlling for amount of social media use. This result is at least suggestive of the negative online influencer culture. We also find, as has been found in other studies, that users who use the internet to create new connections are less likely to feel isolated (**H2**). However, this does not seem to extend to those that use the internet to facilitate offline meetups. Finally, we find that the impact of the internet and social media vary significantly according to the level of urbanicity (**H3**) - this matches with previous findings that suggest that there is significant heterogeneity in the impact of social media on social isloation.

@@

Firstly, we found no evidence that the internet could serve as bridge to build connections for those living outside of cities. The optimistic view that internet use could lower the costs of making connections for those who have less social opportunity, those living in less populated areas, and as a result, help them to feel less isolated. We, in fact, found the opposite. Internet communication had either less positive effect or more negative effect on feelings of social isolation for those living in less urban spaces. This finding lends support to the FOMO channel - viewing lots of content from influencers in a rural area compounds feelings of isolation.

Additionally, given the robustly negative relationship between social media use and social isolation even in the context of China's highly policed internet, our results cast some doubt on theories that suggest it is the negative content, such as cyberbullying or violent/graphic posts, that drives the relationship. Finally, our results also further suggest more work needs to be done in identifying important hetergeneous effects - in our results, those living in a big city experienced almost no change in social isolation as social media use increased while the relationship becomes increasingly negative by ruralness. This specific heterogeneous effect may be unique to China but the strength of it suggests that important inter-personal factors may be relevant in other contexts.

>>>>

-   **H**$1_a$: social media use should have a positive connection to social isolation due to strengthened government control over online content

-   **H**$1_b$: social media use should have a negative connection to social isolation due to the dominance of influencer culture in China, moderated by age.

-   **H**$2$: those who are able to use social media to make new friends will have lower social isolation scores than those who do not

-   **H**$3$: the urban/rural divide will produce differential impacts in the above hypotheses on the relationship between social media use and social isolation

@@

We measured general social media use with an index based on the following three items: 1) About how many hours a day would you estimate you spend using only social media? Social media means applications like Weibo, QQ, Renren, Kaixin001, Douban, WeChat or other sites and services that allow users to interact with each other. (0-1, 1-2, 2-3, 3-4, 4-5, 5-6, 6-7, 7-8, 8-9, or More than 9), 2) Do you check email, read websites, and use social media (social media means applications like Weibo, QQ, Renren, Kaixin001, Douban, WeChat or other sites and services that allow users to interact with each other) more than you did five years ago? (Yes, No), and 3) How often do you read news stories about political events that have been posted on social media (social media means applications like Weibo, QQ, Renren, Kaixin001, Douban, WeChat or other sites and services that allow users to interact with each other)? (More than once a day, Everyday, Three-to-five days per week, One-to-two days per week, Less often, Never). Each was recoded so that higher values represented more social media use, rescaled to range from 0 through 1, added together, and then again, rescaled to range from 0 through 1 maintaining all original intervals.The Chronbach's alpha was relatively low for these three items ($\alpha = 0.31$), but we proceeded with the construction of this index because, first, the face validity is high. Simply, those who claim to use social media generally more often, claim to read political news more often, and claim to use it more than they did in the past, are certainly likely to be more frequent social media users than those who do not do all these things.Second, we confirmed that these three items were all positively correlated with each other (p \< 0.0001 for all pairwise relationships). The distribution of this index is presented in @fig-sm-use-dist. The distribution is relatively normal and tightly centered around the mean 0.63 with a standard deviation of 0.17, but there is a left skew suggesting that a portion of the population is not that social media active, but clearly the bulk of the population are.

@@

Finally, while not a key variable of interest, we also include a dummy variable for Chinese Communist Party (CCP) membership. Previous research on party members suggest that party members may volunteer for social causes at a higher rate than non party members [@dickson2014]. This volunteering behavior may contribute to a lower level of feelings of social isolation and therefore it is of interest to estimate any possible effects of party member status. Additionally, a long line of research has noted that party members earn more than non-party members [@dickson2000; @yan2019]. If party membership affects social isolation and also affects earnings, controlling for CCP membership is important to prevent a biased estimate of the relationship between earnings and social isolation.

Our modeling strategy that follows is straightforward. We begin by modeling each of our internet communication independent variables as function of socio-economic status (SES), gender, CCP membership status, and age to create an individual profile of each type of digital user (see the Online Appendix for the operationalization of these variables). This provides some context for the models that follow which are intended to test our primary hypotheses, that is that internet communication can deter social isolation and that this is particularly true for those who do not have the convenience of the high social interaction opportunity provided by urban living. For those we begin by fitting an additive model of our index of social isolationism as function of each of our primary independent variables while also controlling for the same variables included in our profile models. Finally, we refit three models with the same specification but introduce an interactive term between each of our internet communication indicators, respectively, with our measure of urbanicity. This allows us to test whether the observed additive effects are stronger for those living outside cities.

@@

In terms of our original hypotheses, we find support for the idea that China's social media environment is not particularly unique with respect to generating feelings of social isolation (**H1b**). Whether or not this is due to the influencer culture or other factors, we cannot say with our data, however age is a consistently negative predictor of social isolation even when controlling for amount of social media use. This result is at least suggestive of the negative online influencer culture. We also find, as has been found in other studies, that users who use the internet to create new connections are less likely to feel isolated (**H2**). However, this does not seem to extend to those that use the internet to facilitate offline meetups. Finally, we find that the impact of the internet and social media vary significantly according to the level of urbanicity (**H3**) - this matches with previous findings that suggest that there is significant heterogeneity in the impact of social media on social isolation.

@@

Firstly, we found no evidence that the internet could serve as bridge to build connections for those living outside of cities. The optimistic view that internet use could lower the costs of making connections for those who have less social opportunity, those living in less populated areas, and as a result, help them to feel less isolated. We, in fact, found the opposite. Internet communication had either less positive effect or more negative effect on feelings of social isolation for those living in less urban spaces. This finding lends support to the FOMO channel - viewing lots of content from influencers in a rural area compounds feelings of isolation. The finding by @henning-smith2018 and others that rural respondents are more socially isolated seems, at least in China, to be exacerbated by social media.

Additionally, given the robustly negative relationship between social media use and social isolation even in the context of China's highly policed internet, our results cast some doubt on theories that suggest it is the negative content, such as cyberbullying or violent/graphic posts, that drives the relationship. While @ademiluyi2022 finds that cyberbulling and has a negative short term relationship with social exclusion our finding that there is a negative relationship between social media use and social isolation even in places like China that actively attempt to limit this type of online behavior suggest that the relationship we observe has deeper causes. More generally, China's internet governance strategy of limiting negative online experience does not seem to have dampened the relationship between social media use and social isolation, indicating to us that other proposed mechanisms are likely driving the observed relationship.

Finally, our results also further suggest more work needs to be done in identifying important heterogeneous effects - in our results, those living in a big city experienced almost no change in social isolation as social media use increased while the relationship becomes increasingly negative by ruralness. This specific heterogeneous effect may be unique to China but the strength of it suggests that important inter-personal factors may be relevant in other contexts. Importantly, given our survey results are from 2015, our finding of important heterogeneous effects suggest that these effects did not begin around the time of recent scholarly interest in the topic [@beyens2020; @ostic2021] but rather have been an important feature of the relationship between social media and social isolation since the beginning of social media use.

### 2024-08-19 tandf/Social-Isolation-in-China-second-revision.tex
Review 2 update

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Lonely and online: an investigation of social media use and social isolation in China},
  pdfkeywords={Social Isolation, Social Media, Digital
Communication, China},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Lonely and online: an investigation of social media use and
social isolation in China}
\author{}
\date{2024-06-10}

\begin{document}
\maketitle
\begin{abstract}
Research on the link between social media and well being has lead to
ambiguous findings - some research has found a positive, negative, and
mixed relationship between the two. We employ the case of China, in
which social media is controlled in important ways, as a useful
alternative case to gain leverage on these ambiguities. Using data from
a nationwide survey of Chinese citizens (n=2292), we principally find
that individual, heterogeneous effects are an important driver of the
relationship between levels of social media use and reported well-being.
Additionally, our results lend support to a FOMO (fear of missing out)
mechanism as an important driver of the relationship while not finding
support for the mechanism of exposure to negative content on social
media as a predictor of well-being. Finally, our findings suggest that
institutional context may not play an important role in the social media
- social isolation relationship.
\end{abstract}

@@

\begin{itemize}
\item
  \textbf{H\(1_a\)}: social media use should have a positive connection
  to social isolation due to strengthened government control over online
  content
\item
  \textbf{H\(1_b\)}: social media use should have a negative connection
  to social isolation due to the dominance of influencer culture in
  China, moderated by age.
\item
  \textbf{H\(2\)}: those who are able to use social media to make new
  friends will have lower social isolation scores than those who do not
\item
  \textbf{H\(3\):} the urban/rural divide will produce differential
  impacts in the above hypotheses on the relationship between social
  media use and social isolation
\end{itemize}

@@

\includegraphics{Social-Isolation-in-China-jg-revised_files/figure-latex/fig-social-iso-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-jg-revised_files/figure-latex/fig-var-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-jg-revised_files/figure-latex/fig-sm-use-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-jg-revised_files/figure-latex/fig-urban-dist-1.pdf}

@@

Our modeling strategy that follows is straightforward. We begin by
modeling each of our internet communication independent variables as
function of socio-economic status (SES), gender, Chinese Communist Party
membership (CCP), and age to create an individual profile of each type
of digital user (see the Online Appendix for the operationalization of
these variables). This provides some context for the models that follow
which are intended to test our primary hypotheses, that is that internet
communication can deter social isolation and that this is particularly
true for those who do not have the convenience of the high social
interaction opportunity provided by urban living. For those we begin by
fitting an additive model of our index of social isolationism as
function of each of our primary independent variables while also
controlling for the same variables included in our profile models.
Finally, we refit three models with the same specification but introduce
an interactive term between each of our internet communication
indicators, respectively, with our measure of urbanicity. This allows us
to test whether the observed additive effects are stronger for those
living outside cities.

@@

\includegraphics{Social-Isolation-in-China-jg-revised_files/figure-latex/fig-social-isolation-interactions-1.pdf}

@@

In terms of our original hypotheses, we find support for the idea that
China's social media environment is not particularly unique with respect
to generating feelings of social isolation (\textbf{H1b}). Whether or
not this is due to the influencer culture or other factors, we cannot
say with our data, however age is a consistently negative predictor of
social isolation even when controlling for amount of social media use.
This result is at least suggestive of the negative online influencer
culture. We also find, as has been found in other studies, that users
who use the internet to create new connections are less likely to feel
isolated (\textbf{H2}). However, this does not seem to extend to those
that use the internet to facilitate offline meetups. Finally, we find
that the impact of the internet and social media vary significantly
according to the level of urbanicity (\textbf{H3}) - this matches with
previous findings that suggest that there is significant heterogeneity
in the impact of social media on social isloation.

@@

Firstly, we found no evidence that the internet could serve as bridge to
build connections for those living outside of cities. The optimistic
view that internet use could lower the costs of making connections for
those who have less social opportunity, those living in less populated
areas, and as a result, help them to feel less isolated. We, in fact,
found the opposite. Internet communication had either less positive
effect or more negative effect on feelings of social isolation for those
living in less urban spaces. This finding lends support to the FOMO
channel - viewing lots of content from influencers in a rural area
compounds feelings of isolation.

Additionally, given the robustly negative relationship between social
media use and social isolation even in the context of China's highly
policed internet, our results cast some doubt on theories that suggest
it is the negative content, such as cyberbullying or violent/graphic
posts, that drives the relationship. Finally, our results also further
suggest more work needs to be done in identifying important hetergeneous
effects - in our results, those living in a big city experienced almost
no change in social isolation as social media use increased while the
relationship becomes increasingly negative by ruralness. This specific
heterogeneous effect may be unique to China but the strength of it
suggests that important inter-personal factors may be relevant in other
contexts.

@@

@@

@@

\includegraphics{Social-Isolation-in-China-jg-revised_files/figure-latex/fig-resids-human-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-jg-revised_files/figure-latex/fig-resids-online-offline-1.pdf}

>>>>

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Lonely and online: an investigation of social media use and social isolation in China},
  pdfauthor={Jason Gainous; Andrew W. MacDonald},
  pdfkeywords={Social Isolation, Social Media, Digital
Communication, China},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\begin{document}
\title{Lonely and online: an investigation of social media use and
social isolation in China}

\author{\name{Jason Gainous\textsuperscript{a} and Andrew W.
MacDonald\textsuperscript{b}\thanks{CONTACT Andrew W.
MacDonald. Email: andrew.macdonald@dukekunshan.edu.cn}}
\affil{\textsuperscript{a}Department of Mass Communication, University
of Sharjah, Sharjah, United Arab Emirates;
\textsuperscript{b}Division of Social Sciences, Duke Kunshan
University, Kunshan, China}}

\maketitle

\begin{abstract}
Research on the link between social media and well being has lead to
ambiguous findings - some research has found a positive, negative, and
mixed relationship between the two. We employ the case of China, in
which social media is controlled in important ways, as a useful
alternative case to gain leverage on these ambiguities. Using data from
a nationwide survey of Chinese citizens (n=2292), we principally find
that individual, heterogeneous effects are an important driver of the
relationship between levels of social media use and reported well-being.
Additionally, our results lend support to a FOMO (fear of missing out)
mechanism as an important driver of the relationship while not finding
support for the mechanism of exposure to negative content on social
media as a predictor of well-being. Finally, our findings suggest that
institutional context may not play an important role in the social media
- social isolation relationship.
\end{abstract}

\begin{keywords}
    Social Isolation, Social Media, Digital Communication, 
    China
\end{keywords}
\newpage{}

@@

\begin{itemize}
\item
  \textbf{H}\(1_a\): social media use should have a positive connection
  to social isolation due to strengthened government control over online
  content
\item
  \textbf{H}\(1_b\): social media use should have a negative connection
  to social isolation due to the dominance of influencer culture in
  China, moderated by age.
\item
  \textbf{H}\(2\): those who are able to use social media to make new
  friends will have lower social isolation scores than those who do not
\item
  \textbf{H}\(3\): the urban/rural divide will produce differential
  impacts in the above hypotheses on the relationship between social
  media use and social isolation
\end{itemize}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-social-iso-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-var-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-sm-use-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-urban-dist-1.pdf}

@@

Finally, while not a key variable of interest, we also include a dummy
variable for Chinese Communist Party (CCP) membership. Previous research
on party members suggest that party members may volunteer for social
causes at a higher rate than non party members (Dickson, 2014). This
volunteering behavior may contribute to a lower level of feelings of
social isolation and therefore it is of interest to estimate any
possible effects of party member status. Additionally, a long line of
research has noted that party members earn more than non-party members
(Dickson \& Rublee, 2000; Yan, 2019). If party membership affects social
isolation and also affects earnings, controlling for CCP membership is
important to prevent a biased estimate of the relationship between
earnings and social isolation.

Our modeling strategy that follows is straightforward. We begin by
modeling each of our internet communication independent variables as
function of socio-economic status (SES), gender, CCP membership status,
and age to create an individual profile of each type of digital user
(see the Online Appendix for the operationalization of these variables).
This provides some context for the models that follow which are intended
to test our primary hypotheses, that is that internet communication can
deter social isolation and that this is particularly true for those who
do not have the convenience of the high social interaction opportunity
provided by urban living. For those we begin by fitting an additive
model of our index of social isolationism as function of each of our
primary independent variables while also controlling for the same
variables included in our profile models. Finally, we refit three models
with the same specification but introduce an interactive term between
each of our internet communication indicators, respectively, with our
measure of urbanicity. This allows us to test whether the observed
additive effects are stronger for those living outside cities.

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-social-isolation-interactions-1.pdf}

@@

In terms of our original hypotheses, we find support for the idea that
China's social media environment is not particularly unique with respect
to generating feelings of social isolation (\textbf{H1b}). Whether or
not this is due to the influencer culture or other factors, we cannot
say with our data, however age is a consistently negative predictor of
social isolation even when controlling for amount of social media use.
This result is at least suggestive of the negative online influencer
culture. We also find, as has been found in other studies, that users
who use the internet to create new connections are less likely to feel
isolated (\textbf{H2}). However, this does not seem to extend to those
that use the internet to facilitate offline meetups. Finally, we find
that the impact of the internet and social media vary significantly
according to the level of urbanicity (\textbf{H3}) - this matches with
previous findings that suggest that there is significant heterogeneity
in the impact of social media on social isolation.

@@

Firstly, we found no evidence that the internet could serve as bridge to
build connections for those living outside of cities. The optimistic
view that internet use could lower the costs of making connections for
those who have less social opportunity, those living in less populated
areas, and as a result, help them to feel less isolated. We, in fact,
found the opposite. Internet communication had either less positive
effect or more negative effect on feelings of social isolation for those
living in less urban spaces. This finding lends support to the FOMO
channel - viewing lots of content from influencers in a rural area
compounds feelings of isolation. The finding by Henning-Smith et al.
(2018) and others that rural respondents are more socially isolated
seems, at least in China, to be exacerbated by social media.

Additionally, given the robustly negative relationship between social
media use and social isolation even in the context of China's highly
policed internet, our results cast some doubt on theories that suggest
it is the negative content, such as cyberbullying or violent/graphic
posts, that drives the relationship. While Ademiluyi et al. (2022) finds
that cyberbulling and has a negative short term relationship with social
exclusion our finding that there is a negative relationship between
social media use and social isolation even in places like China that
actively attempt to limit this type of online behavior suggest that the
relationship we observe has deeper causes. More generally, China's
internet governance strategy of limiting negative online experience does
not seem to have dampened the relationship between social media use and
social isolation, indicating to us that other proposed mechanisms are
likely driving the observed relationship.

Finally, our results also further suggest more work needs to be done in
identifying important heterogeneous effects - in our results, those
living in a big city experienced almost no change in social isolation as
social media use increased while the relationship becomes increasingly
negative by ruralness. This specific heterogeneous effect may be unique
to China but the strength of it suggests that important inter-personal
factors may be relevant in other contexts. Importantly, given our survey
results are from 2015, our finding of important heterogeneous effects
suggest that these effects did not begin around the time of recent
scholarly interest in the topic (Beyens et al., 2020; Ostic et al.,
2021) but rather have been an important feature of the relationship
between social media and social isolation since the beginning of social
media use.

@@

\bibitem[\citeproctext]{ref-dickson2014}
Dickson, B. J. (2014). Who Wants to Be a Communist? Career Incentives
and Mobilized Loyalty in China. \emph{The China Quarterly}, \emph{217},
42--68. \url{https://doi.org/10.1017/S0305741013001434}

\bibitem[\citeproctext]{ref-dickson2000}
Dickson, B. J., \& Rublee, M. R. (2000). Membership Has Its Privileges.
\emph{Comparative Political Studies}, \emph{33}(1), 87--112.
\url{https://doi.org/10.1177/0010414000033001004}

@@

\bibitem[\citeproctext]{ref-yan2019}
Yan, S. (2019). The changing faces and roles of communist party
membership in china: An empirical analysis based on CHIPS 1988, 1995 and
2002. \emph{Journal of Contemporary East Asia Studies}, \emph{8}(1),
99--120. \url{https://doi.org/10.1080/24761028.2019.1633987}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-resids-human-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-resids-online-offline-1.pdf}
